### 2020-12-22 ch03_esmvaltool/esmvaltool.tex
Added section on ESMValTool

In order to answer the research questions of this thesis posed in
\cref{sec:01:research_questions}, a reliable and efficient tool to read,
process and evaluate output from climate model simulations is necessary. A
valuable software that fits these criteria is the \ac{ESMValTool}. The
\ac{ESMValTool} is an open-source community diagnostics and performance metrics
tool for the routine evaluation of \ac{ESM} output, which notably facilitates
the analysis of \ac{CMIP} models (\url{https://www.esmvaltool.org}). For this
reason, all analyses presented in
\cref{ch:04:assessment_climate_metrics,ch:05:paper_ecs,ch:06:paper_gpp} of this
thesis were implemented into the \ac{ESMValTool}. Apart from that, further
substantial contributions to the code base of the \ac{ESMValTool} have been
made that improve the routine evaluation of climate models, which is beneficial
for the entire scientific community. Moreover, this lead to co-authorship in
the scientific documentation of the \ac{ESMValTool}, which is published in four
peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
After a brief overview over the \ac{ESMValTool}'s structure, this chapter
presents all these contributions that are not documented in other chapters of
this thesis.

@@

- overview ESMValTool

- CMOR

- Changes in ESMValCore + table

>>>>

In order to answer the research questions of this thesis posed in
\cref{sec:01:research_questions}, a reliable and efficient tool to read,
process and evaluate climate model output and observational data is necessary.
A valuable software that fits these criteria is the \ac{ESMValTool}. The
\ac{ESMValTool} is an open-source community diagnostics and performance metrics
tool for the routine evaluation of \ac{ESM} output, which notably facilitates
the analysis of \ac{CMIP} models (\url{https://www.esmvaltool.org}). For this
reason, all analyses presented in
\cref{ch:04:assessment_climate_metrics,ch:05:paper_ecs,ch:06:paper_gpp} of this
thesis were implemented into the \ac{ESMValTool}. Apart from that, further
substantial contributions to the code base of the \ac{ESMValTool} have been
made that improve the routine evaluation of climate models, which is beneficial
for the entire scientific community. Moreover, this lead to co-authorship in
the scientific documentation of the \ac{ESMValTool}, which is published in four
peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
After a brief overview over the \ac{ESMValTool}'s structure, this chapter
presents all these contributions that are not documented in other chapters of
this thesis.

@@

As shown in \cref{subsec:02:climate_modeling}, climate models have been
continuously improved and extended over the last decades from the relatively
simple \acp{AOGCM} to the complex state-of-the-art \acp{ESM} that include an
immense number of variables and processes. In \acs{CMIP}6, more modeling
institutes provide data for more versions of these complex models running
simulations for more experiments. Consequently, the data volume of the entire
\acs{CMIP}6 archive is expected to reach up to $80 \unit{PB}$
\autocite{Balaji2018}, which is a vast increase in comparison to the $2
\unit{PB}$ of \acs{CMIP}5. The increasing complexity and data volume of the
climate models pose a major challenge for the evaluation and analysis of the
model output. To address this big data challenge and support the scientific
community, the \ac{ESMValTool} has been developed to provide an
\enquote{open-source, standardized, community-based software package for the
  systematic, efficient, and well-documented analysis of \ac{ESM} results}
\autocite{Righi2020}. Moreover, the \ac{ESMValTool} allows a routine comparison
of single or multiple climate models against predecessor versions and/or
observations.

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch03_esmvaltool/figs/esmvaltool_schematic.pdf}
  \caption{Schematic representation of the \acf{ESMValTool}. First, input data
    is preprocessed by the \acf{ESMValCore}. Second, the preprocessed data is
    read by diagnostic scripts, which create the final output (\eg{} plots).
    The user can specify the desired input datasets, preprocessing operations
    and diagnostic scripts in the recipe, which is a configuration file that
    controls the main workflow of the \acs{ESMValTool}. Taken and modified from
    \textcite{Righi2020}.}
  \label{fig:03:esmvaltool_schematic}
\end{figure}

Since its first release in 2016 \autocite{Eyring2016a}, the \ac{ESMValTool} has
been greatly extended and improved. A schematic of the current structure of the
\ac{ESMValTool} is illustrated in \cref{fig:03:esmvaltool_schematic}. The first
major element of this workflow is an extensive preprocessing of the input data.
For this, the Python-based \acf{ESMValCore} provide a set of operators
(\emph{preprocessors}) that are applied to the input data. There are two
classes of preprocessors: non-optional and optional ones. Examples for
non-optional preprocessors are the \ac{I/O} operations \emph{load} (loading
input data from climate models and observations) and \emph{save} (saving the
preprocessed data), which utilize the \ac{netCDF} format (a self-describing,
machine-independent binary file format used for the storage of array-based
scientific data). Further non-optional preprocessors include \emph{checks} and
\emph{fixes} of the input data, which test whether the input data adheres to
standards given by the \ac{CMOR} format and fixes the data if necessary. The
\ac{CMOR} format ensures that the output from the many different modeling
institutes within a generation of \ac{CMIP} follows common standards. The
optional preprocessors include commonly used operations on spatiotemporal
datasets, like spatial and temporal \emph{subsetting}, the calculation of
\emph{spatial and temporal statistics}, horizontal and vertical
\emph{interpolation}, land/sea/ice \emph{masking}, \emph{unit conversion}, or
the calculation of \emph{multi-model statistics}. A further example is the
\emph{variable derivation}, which can be used to derive non-\ac{CMOR} variables
from the input data. The aim of the preprocessor is to facilitate the routine
evaluation of climate models by providing a set of commonly used data
operations. To reduce computation times, \ac{ESMValCore} allows parallel
preprocessing of different datasets.

The second main element of the \ac{ESMValTool} workflow is the calculation of
\emph{diagnostics}. This task is performed by the diagnostic scripts, which can
be written in multiple programming languages. Currently, the languages Python,
NCL, R and Julia are supported. The diagnostic scripts, which contain the code
that runs the actual scientific evaluation, read the preprocessed datasets from
\ac{ESMValCore} and use these to create the final output of the tool. Apart
from \ac{netCDF} files and plots, this includes log files with provenance
information that ensure the reproducibility and transparency of the results.
The whole workflow of the \ac{ESMValTool} is controlled with the \emph{recipe},
which is a configuration file in which the user can specify the desired input
datasets, preprocessing operations and diagnostic scripts. Implementing a new
scientific assessment into the \ac{ESMValTool} usually includes writing a new
recipe and one or more diagnostic scripts. A third major element of the
\ac{ESMValTool} is the \emph{\ac{CMOR}ization} (\ie{} the reformatting of data
so it adheres to \ac{CMOR} standards) of raw observational datasets (not shown
in \cref{fig:03:esmvaltool_schematic}). This ensures that the \ac{ESMValTool}
can process arbitrary observational datasets, which can for example be used to
assess the skill of climate model simulations. More details on the structure of
the \ac{ESMValTool} are given in its extensive documentation
(\url{https://esmvaltool.readthedocs.io}).

To ensure a rapid evaluation of the \acs{CMIP}6 models, the \ac{ESMValTool} is
fully integrated into the infrastructure of the \ac{ESGF}, which provides the
\ac{CMIP} model output for the general public \autocite{Eyring2016b}. As soon
as new model data is published on the \ac{ESGF} servers, it can be accessed
with the \ac{ESMValTool} and analyzed. This instantaneous evaluation of the
\ac{CMIP} models is urgently needed since there is a growing dependency on
\ac{CMIP} products by a broad research community and by national and
international climate assessments. For this reason, the \ac{ESMValTool} is for
example used to evaluate climate model output and observational data in several
chapters of the upcoming Sixth \ac{AR} of the \ac{IPCC}.

The \ac{ESMValTool} is developed open-source on GitHub
(\url{https://github.com/ESMValGroup}). It is released under the Apache
License, version 2.0. The source code of the latest released version of the
\ac{ESMValCore} package, which includes the core functionalities of the
\ac{ESMValTool}, is publicly available at Zenodo \autocite{Andela2020a}.
Similarly, the source code of the latest released version of the
\ac{ESMValTool} package, which includes the recipe and diagnostic scripts that
can for example be used to reproduce scientific assessments, is also available
at Zenodo \autocite{Andela2020}.

\section{Contributions to \acs{ESMValCore}}
\label{sec:03:contributions_to_esmvalcore}

- Changes in \ac{ESMValCore} + table

\section{Contributions to \acs{ESMValTool}}
\label{sec:03:contributions_to_esmvaltool}

### 2020-12-22 end_acronyms/acronyms.tex
Added first part of carbon cycle section
Finished section of ESMs
Added climate feedback figure and adapted margins
Added subsection on mathematical framework for feedbacks
Added subsection on ECS
Started section on techniques for reducing uncertainties in climate model projections
Added section description for section on reducing uncertainties
Added section on emergent constraints
Added subsection on Knutti weighting
Added section on ESMValTool

\begingroup
\setstretch{0.5}
\begin{acronym}
  \acro{AOGCM}{Atmosphere-Ocean General Circulation Model}
  \acro{AR}{Assessment Report}
  \acro{C4MIP}{Coupled Climate-Carbon Cycle Model Intercomparison Project}
  \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
  \acro{CMIP}{Coupled Model Intercomparison Project}
  \acro{CO2}[CO\textsubscript{2}]{carbon dioxide}
  \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
  \acro{DLR}{%
    Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace Center)}
  \acro{ECS}{Equilibrium Climate Sensitivity}
  \acro{ESM}{Earth System Model}
  \acro{ESMValTool}{Earth System Model Evaluation Tool}
  \acro{GHG}{Greenhouse Gas}
  \acro{GPP}{Gross Primary Productivity}
  \acro{IAM}{Integrated Assessment Models}
  \acro{IPA}{%
    Institut für Physik der Atmosphäre (Institute of Atmospheric Physics)}
  \acro{IPCC}{Intergovernmental Panel on Climate Change}
  \acro{MIP}{Model Intercomparison Project}
  \acro{ML}{Machine Learning}
  \acro{RCP}{Representative Concentration Pathway}
  \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
  \acro{SSP}{Shared Socioeconomic Pathway}
  \acro{TCR}{Transient Climate Response}
  \acro{WCRP}{World Climate Research Programme}
  \acro{WGCM}{Working Group on Coupled Modelling}
\end{acronym}
\endgroup

>>>>

\begingroup
\setstretch{0.5}
\begin{acronym}
  \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
  \acro{AOGCM}{Atmosphere-Ocean General Circulation Model}
  \acro{AR}{Assessment Report}
  \acro{C4MIP}{Coupled Climate-Carbon Cycle Model Intercomparison Project}
  \acro{CESM}{Community Earth System Model}
  \acro{CMIP}{Coupled Model Intercomparison Project}
  \acro{CH4}[CH\textsubscript{4}]{methane}
  \acro{CMOR}{Climate Model Output Rewriter}
  \acro{CO}{carbon monoxide}
  \acro{CO2}[CO\textsubscript{2}]{carbon dioxide}
  \acro{CV}{cross-validation}
  \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
  \acro{DLR}{%
    Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace Center)}
  \acro{ECS}{Effective Climate Sensitivity}
  \acro{ENSO}{El Niño-Southern Oscillation}
  \acro{ESGF}{Earth System Grid Federation}
  \acro{ESM}{Earth System Model}
  \acro{ESMValCore}{core functionalities of the ESMValTool}
  \acro{ESMValTool}{Earth System Model Evaluation Tool}
  \acro{GHG}{Greenhouse Gas}
  \acro{GPP}{Gross Primary Productivity}
  \acro{GSAT}{global mean near-surface air temperature}
  \acro{GtC}{gigatonnes of carbon}
  \acro{I/O}{Input/Output}
  \acro{IAM}{Integrated Assessment Models}
  \acro{IPA}{%
    Institut für Physik der Atmosphäre (Institute of Atmospheric Physics)}
  \acro{IPCC}{Intergovernmental Panel on Climate Change}
  \acro{LongRunMIP}{Long Run Model Intercomparison Project}
  \acro{MDER}{Multiple Diagnostic Ensemble Regression}
  \acro{MIP}{Model Intercomparison Project}
  \acro{ML}{Machine Learning}
  \acro{MMM}{multi-model mean}
  \acro{NAO}{North Atlantic Oscillation}
  \acro{NCAR}{National Center for Atmospheric Research}
  \acro{netCDF}{Network Common Data Form}
  \acro{PDF}{probability density function}
  \acro{ppm}{parts per million}
  \acro{RCP}{Representative Concentration Pathway}
  \acro{RMSE}{root-mean-square error}
  \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
  \acro{SSP}{Shared Socioeconomic Pathway}
  \acro{TCR}{Transient Climate Response}
  \acro{TOA}{top of the atmosphere}
  \acro{WCRP}{World Climate Research Programme}
  \acro{WGCM}{Working Group on Coupled Modelling}
\end{acronym}
\endgroup
